Seismic Time Series Analysis\\Statistics for Spatio-Temporal Data\\\textit{-- project draft --}

### 2020-06-10 draft.tex
complete all but conclusions

We will not directly compare the two models, since data from the two stations have different orders of magnitude.

@@

In this case, the employment of \textit{auto.arima} stepwise methodology$^{\ref{footnote:auto.arima}}$ highlights the first model improved with one non-seasonal MA parameter more. This is confirmed by the ACF/PACF of its residuals.
The analysis of ACF/PACF plots of the residuals of the second model do not lead to any improvement.
The ACF and PACF plots of both models are similar to the ones from the chosen model for the Uccle station (\figurename~\ref{fig:arima-24}), and we can notice the same 24 hours remaining periodicity.

The Box-Pierce test statistics for both models accept the independence hypothesis between residuals. The two models have similar AICc values, and similar MAPE on the training set ($\approx6\%$). Thus, we can assess the forecasts for both models in \figurename~\ref{fig:mems:arima}.

@@

Since checking the ACF and PACF plots for the original data could lead to a model with the same order, we fitted it for comparison. The MAPEs are higher but comparable, $7\%$ on the test set and $19\%$ on the train set, the Box-Pierce test refutes the independence between residuals, and $88\%/94\%$ of test data fits inside the $80\%$ and $95\%$ CI. Finally, it shows a variability that tends to admit lower values than the training data shows, including negative ones, while having a lower acceptance for higher values. We omit plots related to this model, since such results confirm the usage of the already selected model on the transformed data.

@@

\subsubsection{Seasonalities divergence}
We will highlight the confidence bands for an estimated error of 67\%. This interval estimation lies on the assumption that values for each single hour in the hourly/weekly seasonality is Normally distributed, a slightly different assumption that is not necessarily refuted by the already seen Q-Q plots.

\begin{figure}[h]
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/seasonalities/before-after-compare-1"}
		\caption{Uccle station}
		\label{fig:season.lockdown:uccs}
	\end{subfigure}
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/before-after-compare-1"}
		\caption{Membach station}
		\label{fig:season.lockdown:mems}
	\end{subfigure}
	\caption{Seasonalities computed before and after the lockdown}{\centering\small Seasonalities extracted together with MA filters, confidence bands $\pm\mathrm{sd}$\\}
	\label{fig:season.lockdown}
\end{figure}
%
As we can see in \figurename~\ref{fig:season.lockdown}, the seasonalities' shapes seem different, in particular comparing the weekly seasonality. The seasonalities from before and after the lockdown are still included in the highlighted confidence bands, apart from a few hours. This is true for both stations.

@@

\subsubsection{Forecast divergence}

\begin{figure}[h]
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-9"}
		\caption{Uccle station}
		\label{fig:forecast.lockdown:uccs}
	\end{subfigure}
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-9"}
		\caption{Membach station}
		\label{fig:forecast.lockdown:mems}
	\end{subfigure}
	\caption{Forecasts from March $2^{nd}$.}{\centering\small In red the actual values, in black the forecasts. March $14^{th}$ at midnight is highlighted.\\}
	\label{fig:forecast.lockdown}
\end{figure}
%
By fitting the model\footnote{Data is different, because we are fitting the model on the original train+test dataset, but if we follow similar steps for both stations we can see that the final model we shall choose is the same. Also, the $\lambda$ parameter for the Box-Cox transformation of Membach data is estimated equally.} on data until March $1^{st}$ we can compute forecasts for 8 weeks ahead, until April $26^{th}$. As we can see in \figurename~\ref{fig:forecast.lockdown}, both stations overestimate starting after the lockdown. Uccle data in particular show a greater overestimation that can be noticed also by looking at the difference between the forecasted and measured values.

Analytically, the MAPE for Uccle doubles with respect to the previous test set ($\approx34\%$), and only 58\% and 78\% of the test data falls inside the confidence intevals (80\% and 95\%). Analysing the Membach station, we find those measures have comparable values to the previous test set: the MAPE remains at $16.6\%$, while 74\% and 90\% of data falls inside the confidence intervals.

@@

\section{Conclusions}
TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO TODO

@@

>>>>

We will not directly compare the two models, since data from the two stations have different orders of magnitude, and that difference should be firstly assessed in order to meaningfully compare the models.

@@

In this case, the employment of \textit{auto.arima} stepwise methodology$^{\ref{footnote:auto.arima}}$ highlights the first model improved with one non-seasonal MA parameter more. This is confirmed by the ACF/PACF plots of its residuals. Their analysis for the second model do not lead to any improvement.
The ACF and PACF plots of both models are similar to the ones from the chosen model for the Uccle station (\figurename~\ref{fig:arima-24}), and we can notice the same 24 hours remaining periodicity.

The Box-Pierce test statistics for both models accept the independence hypothesis between residuals. The two models have similar AIC values, and similar MAPE on the training set ($\approx6\%$). Thus, we can assess the forecasts for both models in \figurename~\ref{fig:mems:arima}.

@@

Since checking the ACF and PACF plots for the original data could lead to a model with the same order, we fitted it for comparison. The MAPEs are higher but comparable, $7\%$ on the test set and $19\%$ on the train set, the Box-Pierce test refutes the independence between residuals, and $88\%/94\%$ of test data fits inside the $80\%$ and $95\%$ CI. Finally, it shows a variability that tends to admit lower values than the training data shows, including negative ones, while having a lower acceptance for higher values. We omit plots related to this model, since such results confirm the choice of the model on the transformed data.

@@

\begin{figure}[h]
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/seasonalities/before-after-compare-1"}
		\caption{Uccle station}
		\label{fig:season.lockdown:uccs}
	\end{subfigure}
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/seasonalities/before-after-compare-1"}
		\caption{Membach station}
		\label{fig:season.lockdown:mems}
	\end{subfigure}
	\caption{Seasonalities computed before and after the lockdown}{\centering\small Seasonalities extracted together with MA filters, confidence bands $\pm\mathrm{sd}$\\}
	\label{fig:season.lockdown}
\end{figure}
%
\subsubsection{Seasonalities divergence}
We highlight the confidence bands for an estimated error of 67\%. This interval estimation lies on the assumption that values for each single hour in the hourly/weekly seasonality is Normally distributed, a slightly different assumption that is not necessarily refuted by the already seen Q-Q plots.

As we can see in \figurename~\ref{fig:season.lockdown}, the seasonalities' shapes seem different, in particular comparing the weekly seasonality. The seasonalities from before and after the lockdown are still included in the highlighted confidence bands, apart from a few hours. This is true for both stations.

@@

\subsubsection{Forecast divergence}
Since both our models forecast stable values with almost stable variance (\figurename{s}~\ref{fig:arima-forecast:200011}, \ref{fig:mems:arima:200011}), we can forecast another 8 weeks ahead, until April $26^{th}$, and then compare the efficency of the results with the original test set. These 8 weeks comprise 6 weeks after March $14^{th}$, first day of lockdown in Belgium.

As we can see in \figurename~\ref{fig:forecast.lockdown}, at both stations the models overestimate starting after the lockdown. Uccle data in particular show a greater overestimation, that can also be noticed by looking at the difference between the forecasted and measured values.

\begin{figure}[h]
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/final-model-5"}
		\caption{Uccle station}
		\label{fig:forecast.lockdown:uccs}
	\end{subfigure}
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/final-model-5"}
		\caption{Membach station}
		\label{fig:forecast.lockdown:mems}
	\end{subfigure}
	\caption{Forecasts from March $2^{nd}$.}{\centering\small In red the actual values, in black the forecasts. March $14^{th}$ at midnight is highlighted.\\}
	\label{fig:forecast.lockdown}
\end{figure}
%
We assess analytical values obtained from the new 8 weeks forecast:
\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
	\item the MAPE for Uccle triples with respect to the original test set to $\approx46\%$, and only 39\% and 63\% of the test data falls inside the confidence intevals (80\% and 95\%).
	\item regarding the Membach station forecasts, the MAPE is reduced to $14\%$, while 82\% and 96\% of data falls inside the confidence intervals.
\end{itemize}

Thus, we notice a visible degradation in performance for Uccle station data, and a slight improvement for Membach station's. The slight improvement is most certainly driven by the single peak, with lower values than previous peaks, and by the increased variability for data further from the end of the training set.

@@

\section{Conclusions}
The models we choose for both stations show interesting predictive efficiency,

@@

% comment by prof on some arima plots
% If you're trying to compare what happens before and after lock-down, you should probably assess the forecast for the period without lock-down and then compare with how it performs for forecasting lock-down values. A large difference (forecast performing worse during lock-down) would indicate that the model doesn't work during lock-down because something has changed in the behaviour of the data, thus confirming your hypothesis
